Binocular combination in the autonomic nervous system

### 2025-12-18 manuscript.tex
Edits

\hypertarget{author-summary}{%
\section{Author Summary}\label{author-summary}}

Our pupils constantly adjust in size to control how much light enters the eye: in brighter conditions the pupils will reduce in size, and in dimmer conditions they will increase in size. This process, known as the pupillary light response, is usually taken for granted but is crucial for vision, and is controlled automatically by the body's nervous system. Different types of light-sensing cells in the retina, the photoreceptor cells, contribute to this response: some respond quickly to changes in light, while others maintain responses over longer periods. In our study, we explored how signals from the two eyes are combined to regulate pupil size. Using a new type of visual stimulation that allowed us to selectively target specific classes of photoreceptor cells, we found that the way the two eyes' signals are combined depends on which class of photoreceptor cells is involved. For example, some photoreceptor classes showed cooperation between the eyes, while others showed competition. By revealing these distinct patterns of signal combination, our work provides new insight into how light information is integrated in the pupils.

@@

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation {[}\protect\hyperlink{ref-Kingdom2015}{23}{]}. We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum {[}\protect\hyperlink{ref-Busse2009}{24}{]}. The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis {[}\protect\hyperlink{ref-Spitschan2019}{25}{]} indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration {[}\protect\hyperlink{ref-Mathot2018}{9}{]}. Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work {[}\protect\hyperlink{ref-Baker2024}{26}{]} has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the light flux pathway. Using a contrast discrimination paradigm with spatial modulations of light flux and colour, we found equally strong interocular suppression in all three pathways {[}\protect\hyperlink{ref-Baker2024}{26}{]}. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the light flux and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in our previous work {[}\protect\hyperlink{ref-Baker2024}{26}{]}. Our other recent work {[}\protect\hyperlink{ref-Segala2023}{3}{]} has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations {[}\protect\hyperlink{ref-Baker2017}{27}{]}. These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' {[}\protect\hyperlink{ref-Baker2007}{20}{]}, in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes {[}\protect\hyperlink{ref-Kaestner2019}{28}{]}. Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

Twenty-four participants were recruited for each of the four experiments for a total of ninety-six adult participants (28 male, 68 female), whose ages ranged from 18 to 41. All participants had normal or corrected to normal vision, no known abnormalities of binocular or colour vision, and gave written informed consent. Our procedures were approved by the Ethics Committee of the Department of Psychology at the University of York (identification number 184).

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB, which are thermally stable across time with active cooling according to the manufacturers: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised binocular viewing system. The light engines were operated via a Python interface to their REST API {[}\protect\hyperlink{ref-Martin2022}{29}{]}, which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. A special command was commissioned from LEDMOTIVE to allow two different stimulus files to be launched simultaneously from the two devices. The synchronisation of the spectra from the two devices was tested and showed that they were synchronised to within approximately 3ms.

When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We used the silent substitution technique {[}\protect\hyperlink{ref-Estevez1982}{30}{]} to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them {[}\protect\hyperlink{ref-Shapiro1996}{31},\protect\hyperlink{ref-Spitschan2018}{32}{]}. We calculated silent substitution solutions using the \emph{PySilSub} toolbox {[}\protect\hyperlink{ref-Martin2023}{33}{]}, using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. Each primary was also linearised using a polynomial fit (see Figure \ref{fig:primarylinearity} for details). We used the 10-degree cone fundamentals {[}\protect\hyperlink{ref-Stockman2000}{34}{]}, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in a previous paper {[}\protect\hyperlink{ref-Martin2023}{33}{]}) to calculate \(\alpha\)-opic irradiance.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies {[}\protect\hyperlink{ref-Barrionuevo2016}{5}--\protect\hyperlink{ref-Spitschan2014}{7},\protect\hyperlink{ref-Spitschan2015}{35},\protect\hyperlink{ref-Zele2018}{36}{]}. The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]}: monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean light flux. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany {[}\protect\hyperlink{ref-Kassner2014}{37}{]}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

Our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]} used a temporal frequency of 2Hz for foveal luminance flicker, and recorded EEG data simultaneously with pupillometry. Initial pilot experiments indicated that this frequency was too high to elicit measurable responses when stimulating individual photoreceptor pathways. For all experiments, we therefore used a primary flicker frequency of 0.5Hz, as previous literature showed that this was slow enough to elicit a pupil response from all photoreceptor classes {[}\protect\hyperlink{ref-Spitschan2014}{7}{]}. We also focussed on only recording pupillometry data as this frequency would be too slow to elicit steady-state EEG responses {[}\protect\hyperlink{ref-Norcia2015}{38}{]}.

@@

Silent substitution stimuli were prepared and calibrated for each participant with custom Python software {[}\protect\hyperlink{ref-Martin2023}{33}{]} and Python scripts. Estimates of photoreceptor spectral sensitivities for each participant were constructed from the known photopigment absorbance spectra {[}\protect\hyperlink{ref-Stockman2000}{34}{]}, taking account of the peak axial density of the respective photopigments, as well as lens {[}\protect\hyperlink{ref-Stockman2000}{34},\protect\hyperlink{ref-Pokorny1987}{39},\protect\hyperlink{ref-Stockman1999}{40}{]} and macular pigment density {[}\protect\hyperlink{ref-Stockman1999}{40},\protect\hyperlink{ref-Bone1988}{41}{]}, in accordance with the field size and age-dependant CIEPO06 observer model {[}\protect\hyperlink{ref-CIE2006}{42}{]}. The melanopic and rhodopic action spectra of the 32-year-old standard observer were taken from CIE S 026 {[}\protect\hyperlink{ref-CIE2018}{43}{]} and then adjusted for age-related lens transmittance with a spectral correction function, in line with the standard. Macular pigment correction was not applied to the rhodopic and melanopic action spectra because rods are not present at the fovea and ipRGCs sit above the retinal pigment layer {[}\protect\hyperlink{ref-Trieschmann2007}{44}{]}.

In the light flux experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substitution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measurable pupil modulations, although we can observe that the rods may not be saturated in the L-M and melanopsin-directed stimuli (Figure \ref{fig:spectraplots}j and \ref{fig:spectraplots}l) and could intrude in these two experiments. We also estimated activation of penumbral L and M cones {[}\protect\hyperlink{ref-Barrionuevo2016}{5},\protect\hyperlink{ref-Spitschan2015}{35},\protect\hyperlink{ref-Zele2019}{45}{]}, which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second {[}\protect\hyperlink{ref-Spitschan2015}{35}{]}, and do not typically affect pupil responses {[}\protect\hyperlink{ref-Spitschan2014}{7}{]}.

@@

Before the start of each experiment, participants adjusted the objective planes of the lenses with the help of the experimenter until the stimulus was in focus and they perceived the two pieces of blackout material as one fused disc. Pupil responses to binocular temporal contrast modulations were examined in a factorial design that combined six ocular conditions and five temporal contrast levels: 6, 12, 24, 48 and 96\% of the available dynamic range. This design, similar to that used in our previous studies {[}\protect\hyperlink{ref-Segala2023}{3},\protect\hyperlink{ref-Baker2017}{27},\protect\hyperlink{ref-Baker2020}{46}{]}, was applied in four separate experiments, each with a different mode of photoreceptor stimulation. In the first three conditions, the discs flickered at 0.5 Hz, in either a monocular, binocular or dichoptic arrangement. In the dichoptic condition the non-target eye saw a flickering fixed contrast of 48\% of the available dynamic range. In the remaining three conditions (the cross-frequency conditions) one eye's disc flickered at 0.4 Hz, and the other eye's disc flickered at 0.5 Hz. This included monocular responses at 0.4 Hz, as well as binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 0.5 Hz, mask contrast of 48\% at 0.4 Hz in the other eye) arrangements. We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

Before the start of the L-M experiment, participants completed a luminance nulling perceptual calibration procedure in L-M cone space on an Iiyama \(\textrm{VisionMaster}^{\textrm{TM}}\) Pro 510 display (800 x 600 pixels, 60 Hz refresh rate). During the task, participants were presented with a disc flickering within the L-M cone space (between magenta and cyan). Using a trackball, participants adjusted the angle in cone space to find their subjective isoluminant point, which resulted in changing the flickering intensity of the stimulus until the amplitude of the flicker appeared to be minimised. The result was used to modify the requested contrasts during stimulus preparation so as to account for individual differences affecting perceived illuminance, principally the L:M cone ratio {[}\protect\hyperlink{ref-Carroll2002}{47},\protect\hyperlink{ref-Hofer2005}{48}{]}.

@@

The pupillometry data were analysed using the same method we used in our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]}. The data were converted from mp4 videos to a csv text file using the Pupil Player software {[}\protect\hyperlink{ref-Kassner2014}{37}{]}, which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. We used coherent averaging and at each stage we excluded data points with a Mahalanobis distance exceeding \(D\) = 3 from the complex-valued mean {[}\protect\hyperlink{ref-Baker2021}{49}{]}. For monocular stimulation, we confirmed that the consensual response was equivalent to the response in the stimulated eye.

@@

The models were fit using a hierarchical Bayesian framework implemented in Stan {[}\protect\hyperlink{ref-Carpenter2017}{50}{]}. The data for each photoreceptor type and response frequency was fit separately, for a total of 8 model fits. The prior for the \(\omega\) parameter was Gaussian, with a mean of 1 and standard deviation of 0.5. Priors for the other free parameters were also Gaussian, with mean values based on previous work {[}\protect\hyperlink{ref-Segala2023}{3}{]}. We calculated over \ensuremath{10^{6}} posterior samples, and retained 10\% for plotting.

@@

\leavevmode\hypertarget{ref-Kingdom2015}{}%
23. Kingdom FAA, Libenson L. Dichoptic color saturation mixture: Binocular luminance contrast promotes perceptual averaging. J Vis. 2015;15: 2. doi:\href{https://doi.org/10.1167/15.5.2}{10.1167/15.5.2}

\leavevmode\hypertarget{ref-Busse2009}{}%
24. Busse L, Wade AR, Carandini M. Representation of concurrent stimuli by population activity in visual cortex. Neuron. 2009;64: 931--42. doi:\href{https://doi.org/10.1016/j.neuron.2009.11.004}{10.1016/j.neuron.2009.11.004}

\leavevmode\hypertarget{ref-Spitschan2019}{}%
25. Spitschan M, Cajochen C. Binocular facilitation in light-mediated melatonin suppression? J Pineal Res. 2019;67: e12602. doi:\href{https://doi.org/10.1111/jpi.12602}{10.1111/jpi.12602}

\leavevmode\hypertarget{ref-Baker2024}{}%
26. Baker DH, Hansford KJ, Segala FG, Morsi AY, Huxley RJ, Martin JT, et al. Binocular integration of chromatic and luminance signals. J Vis. 2024;24(12): 7. doi:\href{https://doi.org/10.1167/jov.24.12.7}{10.1167/jov.24.12.7}

\leavevmode\hypertarget{ref-Baker2017}{}%
27. Baker DH, Wade AR. Evidence for an optimal algorithm underlying signal combination in human visual cortex. Cereb Cortex. 2017;27: 254--264. doi:\href{https://doi.org/10.1093/cercor/bhw395}{10.1093/cercor/bhw395}

\leavevmode\hypertarget{ref-Kaestner2019}{}%
28. Kaestner M, Maloney RT, Wailes-Newson KH, Bloj M, Harris JM, Morland AB, et al. Asymmetries between achromatic and chromatic extraction of 3D motion signals. Proc Natl Acad Sci U S A. 2019;116: 13631--13640. doi:\href{https://doi.org/10.1073/pnas.1817202116}{10.1073/pnas.1817202116}

>>>>

@@

Twenty-four participants were recruited for each of the four experiments for a total of ninety-six adult participants (28 male, 68 female), whose ages ranged from 18 to 41. All participants had normal or corrected to normal vision, no known abnormalities of binocular or colour vision, and gave written informed consent. Our procedures were approved by the Ethics Committee of the Department of Psychology at the University of York (identification number 184). Participant recruitment began on 3 March 2023 and concluded on 8 Decemeber 2023.

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB, which are thermally stable across time with active cooling according to the manufacturers: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised binocular viewing system. The light engines were operated via a Python interface to their REST API {[}\protect\hyperlink{ref-Martin2022}{23}{]}, which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. A special command was commissioned from LEDMOTIVE to allow two different stimulus files to be launched simultaneously from the two devices. The synchronisation of the spectra from the two devices was tested and showed that they were synchronised to within approximately 3ms.

When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We used the silent substitution technique {[}\protect\hyperlink{ref-Estevez1982}{24}{]} to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them {[}\protect\hyperlink{ref-Shapiro1996}{25},\protect\hyperlink{ref-Spitschan2018}{26}{]}. We calculated silent substitution solutions using the \emph{PySilSub} toolbox {[}\protect\hyperlink{ref-Martin2023}{27}{]}, using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. Each primary was also linearised using a polynomial fit (see Figure \ref{fig:primarylinearity} for details). We used the 10-degree cone fundamentals {[}\protect\hyperlink{ref-Stockman2000}{28}{]}, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in a previous paper {[}\protect\hyperlink{ref-Martin2023}{27}{]}) to calculate \(\alpha\)-opic irradiance.

@@

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies {[}\protect\hyperlink{ref-Barrionuevo2016}{5}--\protect\hyperlink{ref-Spitschan2014}{7},\protect\hyperlink{ref-Spitschan2015}{29},\protect\hyperlink{ref-Zele2018}{30}{]}. The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]}: monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean light flux. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany {[}\protect\hyperlink{ref-Kassner2014}{31}{]}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

Our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]} used a temporal frequency of 2Hz for foveal luminance flicker, and recorded EEG data simultaneously with pupillometry. Initial pilot experiments indicated that this frequency was too high to elicit measurable responses when stimulating individual photoreceptor pathways. For all experiments, we therefore used a primary flicker frequency of 0.5Hz, as previous literature showed that this was slow enough to elicit a pupil response from all photoreceptor classes {[}\protect\hyperlink{ref-Spitschan2014}{7}{]}. We also focussed on only recording pupillometry data as this frequency would be too slow to elicit steady-state EEG responses {[}\protect\hyperlink{ref-Norcia2015}{32}{]}.

@@

Silent substitution stimuli were prepared and calibrated for each participant with custom Python software {[}\protect\hyperlink{ref-Martin2023}{27}{]} and Python scripts. Estimates of photoreceptor spectral sensitivities for each participant were constructed from the known photopigment absorbance spectra {[}\protect\hyperlink{ref-Stockman2000}{28}{]}, taking account of the peak axial density of the respective photopigments, as well as lens {[}\protect\hyperlink{ref-Stockman2000}{28},\protect\hyperlink{ref-Pokorny1987}{33},\protect\hyperlink{ref-Stockman1999}{34}{]} and macular pigment density {[}\protect\hyperlink{ref-Stockman1999}{34},\protect\hyperlink{ref-Bone1988}{35}{]}, in accordance with the field size and age-dependant CIEPO06 observer model {[}\protect\hyperlink{ref-CIE2006}{36}{]}. The melanopic and rhodopic action spectra of the 32-year-old standard observer were taken from CIE S 026 {[}\protect\hyperlink{ref-CIE2018}{37}{]} and then adjusted for age-related lens transmittance with a spectral correction function, in line with the standard. Macular pigment correction was not applied to the rhodopic and melanopic action spectra because rods are not present at the fovea and ipRGCs sit above the retinal pigment layer {[}\protect\hyperlink{ref-Trieschmann2007}{38}{]}.

In the light flux experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substitution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measurable pupil modulations, although we can observe that the rods may not be saturated in the L-M and melanopsin-directed stimuli (Figure \ref{fig:spectraplots}j and \ref{fig:spectraplots}l) and could intrude in these two experiments. We also estimated activation of penumbral L and M cones {[}\protect\hyperlink{ref-Barrionuevo2016}{5},\protect\hyperlink{ref-Spitschan2015}{29},\protect\hyperlink{ref-Zele2019}{39}{]}, which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second {[}\protect\hyperlink{ref-Spitschan2015}{29}{]}, and do not typically affect pupil responses {[}\protect\hyperlink{ref-Spitschan2014}{7}{]}.

@@

Before the start of each experiment, participants adjusted the objective planes of the lenses with the help of the experimenter until the stimulus was in focus and they perceived the two pieces of blackout material as one fused disc. Pupil responses to binocular temporal contrast modulations were examined in a factorial design that combined six ocular conditions and five temporal contrast levels: 6, 12, 24, 48 and 96\% of the available dynamic range. This design, similar to that used in our previous studies {[}\protect\hyperlink{ref-Segala2023}{3},\protect\hyperlink{ref-Baker2017}{40},\protect\hyperlink{ref-Baker2020}{41}{]}, was applied in four separate experiments, each with a different mode of photoreceptor stimulation. In the first three conditions, the discs flickered at 0.5 Hz, in either a monocular, binocular or dichoptic arrangement. In the dichoptic condition the non-target eye saw a flickering fixed contrast of 48\% of the available dynamic range. In the remaining three conditions (the cross-frequency conditions) one eye's disc flickered at 0.4 Hz, and the other eye's disc flickered at 0.5 Hz. This included monocular responses at 0.4 Hz, as well as binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 0.5 Hz, mask contrast of 48\% at 0.4 Hz in the other eye) arrangements. We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

Before the start of the L-M experiment, participants completed a luminance nulling perceptual calibration procedure in L-M cone space on an Iiyama \(\textrm{VisionMaster}^{\textrm{TM}}\) Pro 510 display (800 x 600 pixels, 60 Hz refresh rate). During the task, participants were presented with a disc flickering within the L-M cone space (between magenta and cyan). Using a trackball, participants adjusted the angle in cone space to find their subjective isoluminant point, which resulted in changing the flickering intensity of the stimulus until the amplitude of the flicker appeared to be minimised. The result was used to modify the requested contrasts during stimulus preparation so as to account for individual differences affecting perceived illuminance, principally the L:M cone ratio {[}\protect\hyperlink{ref-Carroll2002}{42},\protect\hyperlink{ref-Hofer2005}{43}{]}.

@@

The pupillometry data were analysed using the same method we used in our previous study {[}\protect\hyperlink{ref-Segala2023}{3}{]}. The data were converted from mp4 videos to a csv text file using the Pupil Player software {[}\protect\hyperlink{ref-Kassner2014}{31}{]}, which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. We used coherent averaging and at each stage we excluded data points with a Mahalanobis distance exceeding \(D\) = 3 from the complex-valued mean {[}\protect\hyperlink{ref-Baker2021}{44}{]}. For monocular stimulation, we confirmed that the consensual response was equivalent to the response in the stimulated eye.

@@

The models were fit using a hierarchical Bayesian framework implemented in Stan {[}\protect\hyperlink{ref-Carpenter2017}{45}{]}. The data for each photoreceptor type and response frequency was fit separately, for a total of 8 model fits. The prior for the \(\omega\) parameter was Gaussian, with a mean of 1 and standard deviation of 0.5. Priors for the other free parameters were also Gaussian, with mean values based on previous work {[}\protect\hyperlink{ref-Segala2023}{3}{]}. We calculated over \ensuremath{10^{6}} posterior samples, and retained 10\% for plotting.

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation {[}\protect\hyperlink{ref-Kingdom2015}{46}{]}. We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum {[}\protect\hyperlink{ref-Busse2009}{47}{]}. The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis {[}\protect\hyperlink{ref-Spitschan2019}{48}{]} indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration {[}\protect\hyperlink{ref-Mathot2018}{9}{]}. Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work {[}\protect\hyperlink{ref-Baker2024}{49}{]} has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the light flux pathway. Using a contrast discrimination paradigm with spatial modulations of light flux and colour, we found equally strong interocular suppression in all three pathways {[}\protect\hyperlink{ref-Baker2024}{49}{]}. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the light flux and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in our previous work {[}\protect\hyperlink{ref-Baker2024}{49}{]}. Our other recent work {[}\protect\hyperlink{ref-Segala2023}{3}{]} has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations {[}\protect\hyperlink{ref-Baker2017}{40}{]}. These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' {[}\protect\hyperlink{ref-Baker2007}{20}{]}, in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes {[}\protect\hyperlink{ref-Kaestner2019}{50}{]}. Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

@@

\leavevmode\hypertarget{ref-Martin2022}{}%
23. Martin JT, Pinto J, Bulte D, Spitschan M. PyPlr: A versatile, integrated system of hardware and software for researching the human pupillary light reflex. Behavior Research Methods. 2022;54: 2720--2739. doi:\href{https://doi.org/10.3758/s13428-021-01759-3}{10.3758/s13428-021-01759-3}

\leavevmode\hypertarget{ref-Estevez1982}{}%
24. Estévez O, Spekreijse H. The "silent substitution" method in visual research. Vision Research. 1982;22: 681--691. doi:\href{https://doi.org/10.1016/0042-6989(82)90104-3}{10.1016/0042-6989(82)90104-3}

\leavevmode\hypertarget{ref-Shapiro1996}{}%
25. Shapiro AG, Pokorny J, Smith VC. Conerod receptor spaces with illustrations that use CRT phosphor and light-emitting-diode spectra. Journal of the Optical Society of America A. 1996;13: 2319. doi:\href{https://doi.org/10.1364/josaa.13.002319}{10.1364/josaa.13.002319}

\leavevmode\hypertarget{ref-Spitschan2018}{}%
26. Spitschan M, Woelders T. The Method of Silent Substitution for Examining Melanopsin Contributions to Pupil Control. Frontiers in Neurology. 2018;9. doi:\href{https://doi.org/10.3389/fneur.2018.00941}{10.3389/fneur.2018.00941}

\leavevmode\hypertarget{ref-Martin2023}{}%
27. Martin JT, Boynton GM, Baker DH, Wade AR, Spitschan M. PySilSub: An open-source python toolbox for implementing the method of silent substitution in vision and nonvisual photoreception research. Journal of Vision. 2023;23: 10. doi:\href{https://doi.org/10.1167/jov.23.7.10}{10.1167/jov.23.7.10}

\leavevmode\hypertarget{ref-Stockman2000}{}%
28. Stockman A, Sharpe LT. The spectral sensitivities of the middle- and long-wavelength-sensitive cones derived from measurements in observers of known genotype. Vision Res. 2000;40: 1711--37. doi:\href{https://doi.org/10.1016/s0042-6989(00)00021-3}{10.1016/s0042-6989(00)00021-3}
